---
title: "RiDL timing spot checks"
author: "Michael Hallquist"
date: "23 Feb 2022"
output:
  html_document:
    code_folding: show
    df_print: kable
    mathjax: default
    number_sections: no
    theme: spacelab
    toc: yes
    toc_depth: 2
---

```{r setup}
repo_dir <- "~/Data_Analysis/ridl_fmri_analysis"
#repo_dir <- "/proj/mnhallqlab/users/michael/ridl_fmri_analysis"

ridl_dir <- file.path(repo_dir, "data", "momentum")
setwd(repo_dir)
source("code/parse_ridl.R")

matlab_dir <- "/System/Volumes/Data/Applications/MATLAB_R2020b.app/bin"
```

## Read in the data using parser

```{r}
trial_df <- parse_ridl(sub_id = "220256", ridl_dir = ridl_dir, matlab_dir = matlab_dir)
```

## Expected versus observed RT times

```{r}
# spot checks on timing
# expected versus observed RT
calc_rt <- trial_df$choice_time - trial_df$stim_time
rt_diff <- trial_df$reaction_time - calc_rt 
summary(rt_diff) # looks reasonable
hist(rt_diff)
```

## Expected versus observed feedback onset times

Most times are reasonable (< .2). There is a right tail with values between .2 and 2 seconds

```{r}
# expected versus observed feedback onset
calc_feedback <- trial_df$choice_time + trial_df$feedback_lag # expected feedback onset time
feedback_diff <- trial_df$feedback_time - calc_feedback

summary(feedback_diff)
hist(feedback_diff)
```

## Expected versus observed onset times of next trial

```{r}
# expected versus observed onset of next trial
onset_diff <- trial_df$next_start - (trial_df$feedback_time + 1.0) # 1.0 is fixed duration of feedback event
big_delays <- which(onset_diff > 2)
summary(onset_diff)
hist(onset_diff)
```

## Delays after every 10 trials

Should be 2.5s break after every 10

```{r}
# Every 10 trials, the participant gets a 2.5s break
expected_delays <- which(trial_df$trial %% 10 == 0)
table(abs(onset_diff[expected_delays] - 2.5) < .2)
```

There is one anomaly with a > 2.5 delay after trial 50. It is explained by the subject being too slow on trial 51, leading to repeat
```{r}
# this seems to be a situation where the subject was too slow on trial 51, which explains the delay on a 10-aligned trial
as.data.frame(trial_df[expected_delays[which(abs(onset_diff[expected_delays] - 2.5) > .2)],])
```

## Unexplained delays

These all appear to reflect trial repeats due to slowness
```{r}
# these trials are not on the every-10 trials, but have > 2s delays
weird_delays <- setdiff(big_delays, expected_delays)

# all of the unexpected delays occur on trials where the next trial was too slow
# thus, all of these delays are explained by the repetition of the trial (and this implies that the times in the file are the final presentation)
trial_df$too_slow_lead[weird_delays]

#trial_df %>% filter(times_too_slow > 0) %>% View()
```

```{r, eval = FALSE, include = FALSE}
# this code was from the initial checks before the 1.0s adjustment

# about 40% of trials do have a ~1.0s difference
prop.table(table(abs(onset_diff - 1) < .1))

# 40% of trials have ~ 0s difference
prop.table(table(onset_diff < .1))

# 17% have differences > 1.2s
prop.table(table(onset_diff > 1.2))

# old: this is explained by accrej feedback_time being onset and newlearn being offset. Fixed in parse_ridl.R by subtracting 1.0 in newlearn
# there is some dependency in the offsets -- when onset_diff is 1.0, feedback_diff is often 0.0 and vice versa
# this seems to depend in part on whether it is an accrej or new_learn trial.
data.frame(feedback_diff, onset_diff, trial_df$trial_type) %>% head(n = 20)
```

## Export data to file

```{r}

# write to file
fwrite(trial_df, file=file.path(repo_dir, "output", "example_trial_data.csv"))


## leftovers
#txt <- data.table::fread("/proj/mnhallqlab/studies/momentum/clpipe/data_onsets/220256/Subject_220256.txt")

```

